Cover letter for ``Two applications\dots''

### 2021-03-06 papers/types2020/revisions/sec3.tex

%\newpage
\subsubsection{FPC-specific code}
\label{ssec:fpc}

The \emph{foundational proof certificates} (FPC) framework was
proposed in \cite{chihani17jar} as a flexible approach to specifying a
range of proof structures in first-order classical and intuitionistic
logics.
%
Such specifications are also executable using a simple logic
programming interpreter.
%
As a result of using logic programming, proof certificates in this
framework are allowed to drop details that could be reconstructed during
the checking phase.
%
For example, substitution instances of quantifiers do not need to be
explicitly described within a certificate since unification within the
logic programming checker is often capable of reconstructing such
substitutions.

In this and the next section, we shall only use a much reduced subset
of the FPC framework.
%
In particular, a proof certificate, denoted by the schematic variable
$\XXi$, is a particular term that is threaded throughout a logic
programming interperter.
%
For example, the inference rules in Figure~\ref{fig:augmented} are
augmented with an additional premise called an \emph{expert predicate}
with are responsible for extracting relevant information from a proof
certificate $\XXi$ as well as constructing continuation certificates,
such as, $\XXi_1$ and $\XXi_2$.
%
For example, the premise $\prodE{\XXi}{\XXi_1}{\XXi_2}{t}$ can be
understood as a procedure for extracting from the certificate $\XXi$ a
substitution term $t$ and two continuation certificates $\XXi_1$ and
$\XXi_2$ for the two premises of this rule.
%
If the certificate $\XXi$ does not contain an explicit substitution
term, the expert predicate can simply return a logic variable which would
denote any term that satisfies subsequent unification problems arising
in completing the check of this certificate.

*** DM Continue here

%
%
%%\todo{MM: Talk about randomized FPC? They mentioned afterwards -am}
In our case here, an FPC is a collection of \lP clauses that provide
the remaining details not supplied in Figures~\ref{fig:augmented}
and~\ref{fig:kernel}: that is, the exact set of constructors for the
type of certificates \lsti{cert} as well as the specification of the expert
predicates listed \emph{ibidem}.
%
The top of Figure~\ref{fig:resources} displays two FPCs,
both of which can be used to describe proofs where we bound
the dimension of  a proof.
%
For example, the first FPC dictates that the query \mbox{\lsti{(check (qheight 5) A)}} is
provable in  the kernel using the clauses in Figures~\ref{fig:kernel}
and~\ref{fig:resources} if and only if the height of that proof is 5
or less.
%
Similarly, the second FPC can be used to bound the total number of
instances of the decide rule in a proof.
%
%In particular, the query \mbox{\lsti{(check (qsize 5 H) B)}}
% \begin{lstlisting}
% sigma H\ (check (qsize 5 H) B)
% \end{lstlisting}
%is provable if and only if the total number of unfoldings of that
%proof is 5 or less.

@@

%\newpage

>>>>

\subsubsection{FPC-specific code}
\label{ssec:fpc}

The \emph{foundational proof certificates} (FPC) framework was
proposed in \cite{chihani17jar} as a flexible approach to specifying a
range of proof structures in first-order classical and intuitionistic
logics.  Such specifications are also executable using a simple logic
programming interpreter.  As a result of using logic programming,
proof certificates in this framework are allowed to lack details that
can be reconstructed during the checking phase.  For example,
substitution instances of quantifiers do not need to be explicitly
described within a certificate since unification within the logic
programming checker is often capable of reconstructing such
substitutions.

In this and the next section, we shall only use a much reduced subset
of the FPC framework.  In particular, a proof certificate, denoted by
the schematic variable $\XXi$, is a particular term that is threaded
throughout a logic programming interperter.  For example, the
inference rules in Figure~\ref{fig:augmented} are augmented with an
additional premise that invokes an \emph{expert predicate} with is
responsible for extracting relevant information from a proof
certificate $\XXi$ as well as constructing continuation certificates,
such as, $\XXi_1$ and $\XXi_2$.  For example, the premise
$\prodE{\XXi}{\XXi_1}{\XXi_2}{t}$ calls the expert for products which
should extract from the certificate $\XXi$ a substitution term $t$ and
two continuation certificates $\XXi_1$ and $\XXi_2$ for the two
premises of this rule.  If the certificate $\XXi$ does not contain an
explicit substitution term, the expert predicate can simply return a
logic variable which would denote any term that satisfies subsequent
unification problems arising in completing the check of this
certificate.

In our case here, an FPC is a collection of \lP clauses that provide
the remaining details not supplied in Figures~\ref{fig:augmented}
and~\ref{fig:kernel}: that is, the exact set of constructors for the
type of certificates \lsti{cert} as well as the specification of the
expert predicates listed \emph{ibidem}.  The top of
Figure~\ref{fig:resources} displays two FPCs, both of which can be
used to describe proofs where we bound the dimension of a proof.  For
example, the first FPC dictates that the query \mbox{\lsti{(check
    (qheight 5) A)}} is provable in the kernel using the clauses in
Figures~\ref{fig:kernel} and~\ref{fig:resources} if and only if the
height of that proof is 5 or less.  Similarly, the second FPC can be
used to bound the total number of instances of the decide rule in a
proof.  (Obviously, such proof certificates do not contain, for
example, substitution terms.)
%
%In particular, the query \mbox{\lsti{(check (qsize 5 H) B)}}
% \begin{lstlisting}
% sigma H\ (check (qsize 5 H) B)
% \end{lstlisting}
%is provable if and only if the total number of unfoldings of that
%proof is 5 or less.

@@

%%\todo{MM: Talk about randomized FPC? They mentioned afterwards -am}
